---
title: "Luke Olsen Stat 433 HW 1"
output: github_document
---

``` r
library(readr)
library(rmarkdown)
WI92 <- read_csv("WI92.csv")
```

```{r}
library(readr)
WI92 <- read_csv("WI92.csv")
WI19 <- read_csv("WI19.csv")
```

```{r}
WI92.new <- data.frame(WI92$STRUCTURE_NUMBER_008, WI92$YEAR_BUILT_027, WI92$DECK_COND_058, WI92$SUPERSTRUCTURE_COND_059, WI92$SUBSTRUCTURE_COND_060, WI92$CHANNEL_COND_061, WI92$CULVERT_COND_062, WI92$TOLL_020, WI92$ADT_029, WI92$ROADWAY_WIDTH_MT_051)
WI19.new <- data.frame(WI19$STRUCTURE_NUMBER_008, WI19$YEAR_BUILT_027, WI19$DECK_COND_058, WI19$SUPERSTRUCTURE_COND_059, WI19$SUBSTRUCTURE_COND_060, WI19$CHANNEL_COND_061, WI19$CULVERT_COND_062, WI19$TOLL_020, WI19$ADT_029, WI19$ROADWAY_WIDTH_MT_051)
table(WI92.new$WI92.TOLL_020)
table(WI19.new$WI19.TOLL_020)
```

**Looking at this, we see that all bridges in Wisconsin fall under code 3, which means that none of the bridges in 1992 and none in 2019 have tolls.**

```{r}
hist(WI92.new$WI92.YEAR_BUILT_027)
hist(WI19.new$WI19.YEAR_BUILT_027)
```

**Looking at the two histograms, we can see between 1990 and 2010, a large amount of new bridges were built; however, the number is declining moving into 2010-2020.**

```{r}
mean(WI92.new$WI92.ADT_029)
mean(WI19.new$WI19.ADT_029)
plot(WI92.new$WI92.ADT_029, type = "l")
plot(WI19.new$WI19.ADT_029, type = "l")
```

**Here we can see the mean average daily traffic has increased by over 1000 from our 1992 dataset to our 2019 dataset, showing an increase of cars on Wisconsin roads. This is also shown in the line plots, as in the 1992 plot, the average daily traffic never reaches over 140,000 and looking at the graph of 2019, the average daily traffic surpasses 150,000 a few times.**

```{r}
mean(WI92.new$WI92.ROADWAY_WIDTH_MT_051[1:13082])
mean(WI19.new$WI19.ROADWAY_WIDTH_MT_051)
```

**By looking at these two means of the average bridge width from 1992 and 2019, respectively, we can see that bridges are becoming wider. This could be due to the increase in the daily traffic, as bridges are having to be built with more lanes in order to control traffic.**
